---
output:
  pdf_document: default
  html_document: default
---
# Association Rules Mining Assingment
#### Knowledge Discovery and Data Analytics I
#### Win Phyo
#### April 3, 2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
rm(list=ls())  # Clear environment
```
Import necessary packages
```{r message=FALSE}
library('arules')
library('backports')
library('zeallot')
library('arulesViz')
library('dplyr')
library('stringr')
library('chron')
```

```{r message=FALSE}
df <- read.csv('OnlineRetail.csv', stringsAsFactors = FALSE)
```
**Data exploration**
```{r}
str(df)
summary(df)
```
**Data cleaning**
```{r}
sum(is.null(df$InvoiceNo))  # Get sum of all records with null InvoiceNo
```
No null InvoiceNo values observed in dataframe

It is observed that there are invoice numbers that begin with the character 'C'.
Drop 'C' character.
```{r}
xrows <- 0
for(i in 1:nrow(df)){
  if(grepl('C', df[i, 'InvoiceNo'])){
    xrows[i] <- i
  }
}

xrows <- xrows[!is.na(xrows)]  # Drop rows with NA; keep only valid row numbers

df <- df[-xrows, ]  # Remove these rows from dataframe
```
Replace spaces in item description field with underscores
```{r}
df$Description <- trimws(df$Description)  # Remove trailing and leading spaces
df$Description <- gsub(" ", "_", df$Description)
```
Convert date field to native R datetime object
```{r}
df$InvoiceDate <- as.Date(df$InvoiceDate, format = "%m/%d/%Y")
```
Filter dataframe for Irish transactions
```{r}
eire <- df[df$Country == 'EIRE', ]
```
Write cleaned Irish dataset to file for ease of use and as a checkpoint
```{r}
write.csv(eire, file = '2021-cleaned-eire.csv', row.names = FALSE)
```
##### Assocation Rule Mining
Import Irish dataset as transactions
```{r}
eire <- read.transactions(
  '2021-cleaned-eire.csv', 
  format = c('single'),
  header = TRUE,
  rm.duplicates = FALSE,
  cols = c('InvoiceNo', 'StockCode'),
  sep = ','
)
```
Check what this object looks like
```{r}
inspect(eire[1:2])
```
Items are grouped by invoice number; each record in this dataset corresponds to 
a particular, unique invoice number.  

**Explore support counts to determine minimum value/cutoff point**
```{r}
itemFrequencyPlot(eire, topN = 15)  # 15 most frequently occuring items
itemFrequencyPlot(eire, support = 0.1)
itemFrequencyPlot(eire, support = 0.08)
```  

There are 13 items that appear in at least 10% of all transactions. There are 21
items that appear in at least 8% of all transactions.  

**Get association rules**  
Let the minimum confidence value be 0.7, and the minimum support value be 0.08
```{r}
eire.rules <- apriori(
  eire, 
  parameter = list(
    confidence = 0.7,
    support = 0.08,
    minlen = 2
  )
)
```
```{r}
plot(
  eire.rules, 
  method = 'matrix',
)
```

Explore rules
```{r}
inspect(eire.rules)
summary(eire.rules)
```
Return only rules with a lift value greater than 5
```{r}
subset.rules = eire.rules[quality(eire.rules)$lift > 5]
inspect(subset.rules)
```
These rules have at least 8% support, 72% confidence, and 5.5 lift.  

There corresponding item descriptions for the relevant stock codes are as 
follows:

| StockCode | Item Description                  |
|-----------|-----------------------------------|
| 22423     | REGENCY_CAKESTAND_3_TIER          |
| 22697     | GREEN_REGENCY_TEACUP_AND_SAUCER   |
| 22698     | PINK_REGENCY_TEACUP_AND_SAUCER	  |
| 22699     | ROSES_REGENCY_TEACUP_AND_SAUCER	  |

3/4 items are teacup and saucer pairs. Transactions that involved the Regency 
Cakestand and Green Regency Teacup and Saucer also included the Roses Regency
Teacup and Saucer 100% of the time. The rules with the highest lift value are 
rules 7, 1, and 2.
















